---
title: "R Notebook"
output: html_notebook
---

```{r,message=FALSE}
library(TCGAbiolinks)
library(survival)
library(SummarizedExperiment)
```

```{r}
load("dataBRCA/BRCA_FPKM.rda")
BRCA_FPKM <- data
rm(data)
gc()
```

```{r}
# Remove FFPEs as suggested
BRCA_FPKM <- BRCA_FPKM[,!BRCA_FPKM$is_ffpe]
BRCA_FPKM <- BRCA_FPKM[,which(BRCA_FPKM$gender == "female")] ## only female data
BRCA_FPKM <- BRCA_FPKM[,BRCA_FPKM$shortLetterCode=="TP"] ## only tumor tissue

# Make vectors of 5 most favorable and unfavorable prognostic genes, according to the original publication. Subset the data to only contain the 'prognostic genes'.
prognogenes <- c("ENSG00000105825", "ENSG00000104998", "ENSG00000132465", "ENSG00000196843", "ENSG00000023892", "ENSG00000120256", "ENSG00000102144", "ENSG00000120265", "ENSG00000150756", "ENSG00000147676")
BRCA_FPKM <- BRCA_FPKM[prognogenes, ]

# Subset the data to only contain data needed to make Kaplan-Meier plots
colData(BRCA_FPKM) <- colData(BRCA_FPKM)[, c('bcr_patient_barcode','age_at_diagnosis', 'days_to_death', 'days_to_last_follow_up', 'vital_status')]

# Adress arbitrary -Inf to allow filtering of data
colData(BRCA_FPKM)[grep("alive", colData(BRCA_FPKM)$vital_status, ignore.case = TRUE), "days_to_death"] <- "-Inf"
colData(BRCA_FPKM)[grep("dead", colData(BRCA_FPKM)$vital_status, ignore.case = TRUE), "days_to_last_follow_up"] <- "-Inf"

# Remove NA's and duplicates
BRCA_FPKM <- BRCA_FPKM[,!is.na(BRCA_FPKM$days_to_death) &
                        !is.na(BRCA_FPKM$days_to_last_follow_up) &
                        !is.na(BRCA_FPKM$age_at_diagnosis) &
                        !duplicated(BRCA_FPKM$bcr_patient_barcode)]

# Make age more readable by diving by 365
BRCA_FPKM$age_at_diagnosis <- BRCA_FPKM$age_at_diagnosis/365
```

Generate survival object
```{r}
ttime <- BRCA_FPKM$days_to_death
status <- ttime > 0
ttime[!status] <- BRCA_FPKM$days_to_last_follow_up[!status]
ttime[which(ttime == -Inf)] <- 0
ttime=as.numeric(ttime)/365
ttime <- Surv(ttime, status)
```

Make Kaplan-Meier curves for all progonostis genes with the gene expression thresholds obtained from the original publication
```{r}
genesthresholddf <- data.frame(prognogenes,
                          c('TFPI2', 'IL27RA', 'JCHAIN', 'ARID5A', 'DEF6',
                            'LRP11', 'PGK1', 'PCMT1', 'FAM173B', 'MAL2'))

genesthresholddf$cutoffs <- c(0.2451166, 3.1645871, 12.6164274, 5.126073,
                              5.342503, 24.645589, 101.816227, 27.257553,
                              7.340998, 126.772433) 
# 'optimal' cut-off values (FPKM) as obtained from the original publication
# Below, we show how these cut-off values were obtained for one of the genes

colnames(genesthresholddf) <- c('ensembl', 'symbol','cutoffs')

for (i in 1:10) {
  pVal = as.numeric(1 - pchisq(abs(survdiff(ttime~as.character(assays(BRCA_FPKM)$'HTSeq - FPKM'[i, ] < genesthresholddf$cutoffs[i]))$chisq),1)) 
  png(paste0("figuresBRCA/KaplanMeier_", genesthresholddf[i,2], ".png")) # Uncomment to save in a Plots folder
  
  plot(survfit(ttime~as.character(assays(BRCA_FPKM)$'HTSeq - FPKM'[i, ] < genesthresholddf$cutoffs[i])),
       col=1:2,
       main=paste("BRCA:  Kaplan-MeierSurvival (p =",format(pVal,digit=3),")"),
       xlab="Years",
       ylab="Survival")
  
legend("topright",
       legend=paste(c("high expression","low expression"),
                    genesthresholddf[i,2]),col=1:2,text.col=1:2,lty=1)
  dev.off()
}
```

Visualize the p-hacking strategy

```{r}
floor <- unname(quantile(assay(BRCA_FPKM)["ENSG00000104998", ], probs = 0.2))
ceiling <- unname(quantile(assay(BRCA_FPKM)["ENSG00000104998", ], probs = 0.8))

# create a vector of "all" possible thresholds
thresholdVector <- seq(floor,ceiling, by = 0.01)

# create a vector for p-values corresponding to these thresholds
pVals <- c()

for (i in 1:length(thresholdVector)) {
  pVal <- as.numeric(1 - pchisq(abs(survdiff(ttime~as.character(assay(BRCA_FPKM)["ENSG00000104998", ] < thresholdVector[i]))$chisq),1))
  
  pVal <- -log10(pVal)
  pVals[i] <- pVal
}

#png("figuresBRCA/p-hacking.png")
plot(thresholdVector,pVals,type="l",main="ILR27A",xlab="FPKM value",ylab="-log10 p-value",ylim=c(0,6),cex.main=0.8,cex.lab=0.8,cex.axis=0.8)
points(x=3.1645,y=max(pVals),col="red",lwd=2,cex=1.5)
#dev.off()
```

Here we see the profile of -log10 transformed p-values for different FPKM cut-off values. The red circle shows the FPKM value obtained in the original publication. We obtain exactly the same 'optimal' cut-off value as the original authors.





